### 2022-11-30 report/sections/abstract.tex

The transformation of the global energy system to reach climate-neutrality requires large investments into new technologies. \dots