### 2022-11-30 report/sections/introduction.tex

\textbf{Literature}

\begin{itemize}
    \item CCUS or CCTUS seem to be covered mainly by IAM community, not energy system modelling. 
    \item \cite{CaptureMapGetSitespecific} Recent Swiss study on CCTS
    \item \cite{weiProposedGlobalLayout2021} Chinese Nature study on global CCS potential: 92 GtCO$_2$ mitigated globally of which 59 Gt will be sequestered. All at cost of 0.12\% of global GDP   
    \item \cite{ToolsGreenTransition} Endravas Capture Map estimates 1.4 Gt capturing potential from point sources in Europe.
    \item SimCCS Optimization software for MIP problems with infrastructure decision \cite{middletonSimCCSOpensourceTool2020}
\end{itemize}

\textbf{Companies and projects}
\begin{itemize}
    \item Potential Carbon Management hubs are evolving Tree Energy, Carbfix, Equinor    
    \item Mission Innovation Hydrogen Valley Platform \cite{H2ValleysMissionInnovation} collects hydrogen flagship projects to facilitate 
    \item European Hydrogen Backbone \cite{gasforclimateEuropeanHydrogenBackbone2022} 
\end{itemize}

\textbf{Legislation}
\begin{itemize}
    \item German climate protection act \cite{KlimaschutzgesetzKlimaneutralitaetBis} targets climate neutrality by 2045, and 65\% CO$_2$ reduction by 2030. (762 Mt CO2-eq emitted in 2021; 438 Mt targeted for 2030) 
    \item Fit-for-55 package by EU targets 55\% reduction by 2030 in the EU
    \item REpowerEU \cite{REPowerEU} targets accelerated energy transition and security. A fast ramp up of a green Hydrogen system should support transition of hard-to-abate sectors. Domestic production should be 10 MT H$_2$ by 2030, same as import. Additional funding for Clean Hydrogen Partnership through Horizon Europe Programme. 
    \item Delegated Act for green H$_2$
    \item 
\end{itemize}

\textbf{Research Questions}

\begin{itemize}
    \item How to optimally supply cities far from renewable wind production centers in throughout the whole year?    
    \item Where is synergy and where competition between H$_2$ and CO$_2$ management?
    \item Where is one technology in advantage?
    \item DAC vs. point source CC?
    \item Where to draw the line between small and large scale CO$_2$ transport?
\end{itemize}

\textbf{Background Research}

\begin{itemize}
    \item Economy of scale? What are the learning curves to expect?
\end{itemize}