### 2022-11-30 report/sections/methods.tex

Energy system modelling in PyPSA-EUR-sec \cite{brownSynergiesSectorCoupling2018} with version