### 2023-02-07 report/sections/methods.tex

Energy system modelling in PyPSA-EUR-sec~\cite{brownSynergiesSectorCoupling2018}