### 2023-02-23 report/sections/methods.tex
add cluster config

Energy system modelling in PyPSA-EUR-sec~\cite{brownSynergiesSectorCoupling2018}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{sequestration_map.pdf}
    \caption{Sequestration map}
    \label{fig:sequestration_map}
\end{figure}

\begin{figure*}[h]
    \centering
    \includegraphics[width=\linewidth]{capacity_map_co2network_1000.pdf}
    \caption{Optimal capacities per sector for a sequestration of 1000 Mt/a.}
    \label{fig:capacity_map_noco2network_1000}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_1000.pdf}
    \caption{Optimal operation per sector for a sequestration of 1000 Mt/a.}
    \label{fig:operation_map_noco2network_1000}
\end{figure*}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{operation_area_carbon_co2network.pdf}
    \caption{Balance of captured CO$_2$ emissions for the optimal operation with CO$_2$ network.}
    \label{fig:operation_area_carbon_co2network}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{operation_area_carbon_noco2network.pdf}
    \caption{Balance of captured CO$_2$ emissions for the optimal operation without CO$_2$ network.}
    \label{fig:operation_area_carbon_noco2network}
\end{figure}