### 2023-06-13 report/submissions/eem/presentation/presentation.tex

\documentclass[12pt, aspectratio=169]{beamer}
\usetheme{metropolis}
\usepackage{graphicx}
\usepackage{xcolor}

\newcommand{\carbon}{\textcolor{teal}{CO$_2$}}
\newcommand{\hydrogen}{H$_2$}

\title{Designing a \carbon{} Network for a Carbon-Neutral European Economy}
\author{\textbf{Fabian Hofmann}, Christoph Tries, Fabian Neumann, Lisa Zeyen, Tom Brown}
\institute{Technical University Berlin}
\date{}
\urlstyle{blue}

\graphicspath{
    {figures},
}

\begin{document}

\begin{frame}[plain]
    \titlepage
\end{frame}

\begin{frame}
    \begin{columns}
        \column{0.4\textwidth}
        European \carbon{} emissions have to decrease \textbf{quickly}.
        \column{0.6\textwidth}
        \includegraphics[width=\linewidth]{emissions-timeline}
    \end{columns}
\end{frame}

\begin{frame}
    \begin{columns}
        \column{0.4\textwidth}
        Particular focus is needed on \textbf{hard-to-abate emissions} \\
        \vspace*{1em}
        \visible<2->{Formulate a strategy for \\ \textbf{carbon management}}
        \visible<3>{
            \begin{itemize}
                \item \carbon{} Capture
                \item \carbon{} Utilization
                \item \carbon{} Transport
                \item \carbon{} Sequestration
            \end{itemize}

        }
        \column{0.6\textwidth}
        \includegraphics[width=\linewidth]{emissions-timeline-alerted}
    \end{columns}
\end{frame}

\section*{Model Setup}

\begin{frame}
    \begin{columns}
        \column{0.55\textwidth}
        Investment optimization with \textbf{PyPSA-Eur}
        \visible<2>{
            \begin{itemize}
                \item European countries (EU+UK+NO+CH)
                \item 90 nodes
                \item 3 hour resolution
                \item electricity, heating, transport, industry \& waste sectors
                \item carbon management technologies (DAC, CC at point sources, Fischer-Trosch, Methanolization, )
            \end{itemize}
        }
        \column{0.5\textwidth}
        \includegraphics[width=\linewidth]{brownfield_capacities}
    \end{columns}
\end{frame}

\begin{frame}[plain]
    \begin{columns}
        \column{0.3\textwidth}
        Perform 2 $\times$ 5 model runs with
        \carbon{}~\textbf{transport network} on/off and 5~different rates of annual \textbf{sequestration}.\\
        \column{0.7\textwidth}
        \includegraphics[width=\linewidth]{sequestration_map.png}
    \end{columns}
\end{frame}

\section*{System Costs}
\begin{frame}[plain]
    % \qquad \textbf{System costs for different sequestration rates}
    \includegraphics[width=\linewidth]{cost_area.pdf}
\end{frame}

\section*{Carbon Capture Technologies}
\begin{frame}
    \qquad \textbf{Captured carbon balance with a CO$_2$ network}
    \begin{columns}
        \column{0.3\textwidth}
        \begin{itemize}
            \item  Fischer-Tropsch fuels are replaced by fossils fuels
            \item \hydrogen{} infrastructure is replaced by CCS
        \end{itemize}
        \column{0.7\textwidth}
        \begin{figure}
            \includegraphics[width=\linewidth]{operation_area_carbon_co2network.pdf}
            \par\footnotesize{*with Carbon Capture}
        \end{figure}
    \end{columns}
\end{frame}

\begin{frame}
    \textbf{Carbon Capture (left) and Utilization \& Storage (right) for 200~Mt$_{\text{seq}}$/a}
    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_200.png}
\end{frame}

\begin{frame}
    \textbf{Carbon Capture (left) and Utilization \& Storage (right) for 400~Mt$_{\text{seq}}$/a}
    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_400.png}
\end{frame}

\begin{frame}
    \textbf{Carbon Capture (left) and Utilization \& Storage (right) for 600~Mt$_{\text{seq}}$/a}
    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_600.png}
\end{frame}

\begin{frame}
    \textbf{Carbon Capture (left) and Utilization \& Storage (right) for 800~Mt$_{\text{seq}}$/a}
    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_800.png}
\end{frame}

\begin{frame}
    \textbf{Carbon Capture (left) and Utilization \& Storage (right) for 1000~Mt$_{\text{seq}}$/a}
    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_1000.png}
\end{frame}

\section*{Conclusion}

\begin{frame}
    \textbf{Conclusion}
    \begin{itemize}
        \pause
        \item \carbon{} transport is relatively cheap and leads to large system cost-benefits
              \pause
        \item Competition between CCU + \hydrogen{} network and Fossils + CCS
              \pause
        \item Optimal investments strongly dependent on sequestration rates
              \pause
        \item CCU is at point-sources in the inland while DAC is at the coasts
              \pause
        \item Sequestration technology is still to be elaborated $\rightarrow$ take results with care
    \end{itemize}
\end{frame}

\end{document}