### 2023-12-30 report/submissions/journal/cover_letter/letter.tex
add cover letter

% FDUletter_example.tex - an example latex file to illustrate FDUletter.cls
%
% Template by Brian Wood (brian.wood@oregonstate.edu).  Please feel free to send suggestions for changes; this template/cls is not exactly elegantly done!
% Modified by Huang Weiran (huangweiran1998@outlook.com) to fit the need of FDU students.
% Based on SCU version and delete the watermark
% SHU version modified by Zhendong Li (zhendong.li2001@outlook.com) to fit SHU students.

\documentclass[12pt]{SHUletter}
\usepackage{tikz}
\usepackage{eurosym}
\usepackage{xcolor}
	\definecolor{tubred}{RGB}{153,0,0}
\usepackage{lipsum}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{eso-pic}
%
% This section is just a bunch of busywork so that the second and following pages read ``Page X of Y''
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\rhead{Page \thepage \hspace{1pt} of \pageref{LastPage}}
%
%
% Set custom font here. Comment this line out if you do not have a Cambria font (originally included with this template) installed; computer modern (or whatever your current default font is) will be substituted.
%
%\setmainfont{[Cambria.ttf]}[BoldFont  = [CambriaBold.ttf], ItalicFont  = [CambriaItalic.ttf], BoldItalicFont = [CambriaBoldItalic.ttf] ]

\newcommand{\watermark}[3]{\AddToShipoutPictureBG{
\parbox[b][\paperheight]{\paperwidth}{
\vfill%
\centering%
\begin{tikzpicture}
    \path (0,0) -- (\paperwidth,\paperheight);
    \node[opacity=.06] at (current page.center)
    {\includegraphics[width=0.55\textwidth]{tublogo.pdf.pdf}};   % sculogo watermark
    \end{tikzpicture}
\vfill}}}

% The material below is a whole big dang thing whose purpose is just to set up a fixed coordinate system for \tikz so that you can put the Department or School address in the upper right-hand side without it moving all around every time you change something in the page.  I think it works.
% Defining a new coordinate system for the page:
%
% --------------------------
% |(-1,1)    (0,1)    (1,1)|
% |                        |
% |(-1,0)    (0,0)    (1,0)|
% |                        |
% |(-1,-1)   (0,-1)  (1,-1)|
% --------------------------
\makeatletter
\def\parsecomma#1,#2\endparsecomma{\def\page@x{#1}\def\page@y{#2}}
\tikzdeclarecoordinatesystem{page}{
	\parsecomma#1\endparsecomma
	\pgfpointanchor{current page}{north east}
	% Save the upper right corner
	\pgf@xc=\pgf@x%
	\pgf@yc=\pgf@y%
	% save the lower left corner
	\pgfpointanchor{current page}{south west}
	\pgf@xb=\pgf@x%
	\pgf@yb=\pgf@y%
	% Transform to the correct placement
	\pgfmathparse{(\pgf@xc-\pgf@xb)/2.*\page@x+(\pgf@xc+\pgf@xb)/2.}
	\expandafter\pgf@x\expandafter=\pgfmathresult pt
	\pgfmathparse{(\pgf@yc-\pgf@yb)/2.*\page@y+(\pgf@yc+\pgf@yb)/2.}
	\expandafter\pgf@y\expandafter=\pgfmathresult pt
}
\makeatother

%%------------------------------------------------------------------------------%%
%

%%%%%%%%%%% Put Recommender(your mentor) Information Here %%%%%%%%%%%
%
\def\name{Fabian Hofmann, Christoph Tries, Fabian Neumann, Elisabeth Zeyen, Tom Brown.
}
%
% List your degree(s), licences, etc. here if you like.
%\def\What{, Your degrees, etc.}
%
% Set the name of your Department or School here
% I honestly don't know why the negative spacing is necessary to get the alignment of the first line correct.  This must be a ``\tikz'' thing.
%%%%%%%%%%%%%%%%%%  School or Department %%%%%%%%%%%%%%%
\def\Where{\hspace{-1.2mm}\textbf{\color{tubred}
	Department of Digital\\Transformation in Energy Systems,\\ TU Berlin
}}

%%%%%%%%%%%%  Additional Contact Information %%%%%%%%%%%
%
% Set your preferred primary contact address here.
\def\Address{Straße des 17. Juni 135
}
%
\def\CityZip{Berlin, 10623 \\
	Germany}
%
% Set your e-mail here
\def\Email{\textbf{\color{tubred}E-mail}: m.hofmann@tu-berlin.de}
%
% Set your preferred contact number here
%\def\TEL{\textbf{\color{tubred}Tel/Fax}: +86 28 85412720}
%
% Set your department or personal website here
\def\URL{\textbf{\color{tubred}URL}: {https://www.ensys.tu-berlin.de}}
%

%%%%%%%%%%%%%%%%  Footer information  %%%%%%%%%%%%%%%%%%
%
%  The next line is for your college, used as a footer.  If you prefer not to have this, just comment out these lines in favor of the line labeled "[[Alternate]]" below
\def\school{\small{
		TU Berlin $\cdot$
		~Department of Digital\\ Transformation in Energy Systems $\cdot$
		~25, Einsteinufer  $\cdot$
		~Berlin, Germany} }
% \def\school{~}  % [[Alternate]]
%

%%%%%%%%%%%%%%%%%%%%%  Signature line  %%%%%%%%%%%%%%%%%%%%%
%
% Set your signature line here.
% One can add a signature image in a PDF file using the following code; this requires a file called "signature_block.pdf" to be installed in the same folder as the .tex file.  The vertical spacing (\vspace) and the scaling will have to be adjusted to get things to look correct for your particular signature image. Alternatively, comment out the following line in favor of the one labeled "[[Alternate]]" if you want to sign a paper copy of the letter.
%
\signature{
	\name
}
%\signature{\name}  % [[Alternate]]
%%------------------------------------------------------------------------------%%

% This block sets up the address on the right-hand side of the header.
%
% The following lines just compile the information you set up into the LaTex letter variable "address" for later use.
%
%The following command "clears out" the default address so that it can be better set using \tikz
\address{}

\def\newaddress{
	\Where\\
	\Address\\
	\CityZip\\
	%\TEL\\
	\Email\\
	\URL
}
%
%%%%%%%%%%%  DATE  %%%%%%%%%%%%%%%%%%%%%%%%%
% If you want a date different from the current date, comment out the next line in favor of the line labeled "[[Alternate]]".  The ``\vspace{10mm}'' just moves the date down a tiny bit so it doesn't interfere with the header.  This can be adjusted to your preference.
%
\date{\vspace{11mm} \today}
%\date{\vspace{10mm} 20 September 2020}  %[[Alternate]]
%
%%%%%%%%%%% Set the subject here if there is one  %%%%
%\subject{Stuff} % optional subject line

\newcommand{\carbon}{CO$_2$}
\newcommand{\hydrogen}{H$_2$}

%%-----------------------------------------------------------------------------------------%%
%
\begin{document}
	%
	%
	%%%%%%%%  The "To" address goes here.
	%
   %Also, you can use this
	% \begin{letter}{
	% 		Target University\\
	% 		Some Addresss\\
	% 		SomeTown, SomeState 					       				  		 ~~SomeZip
	% 	}
 \begin{letter}{Dear Editor,}

		% This line sets up the return address to the right-side of the OSU logo.  The location is set with absolute node addresses using ``\tikz''.  It can still be a bit fussy, and you may need to alter this a little to get things to look right.  The bit that changes the position are the numbers in parentheses ``at (14.2,2.7)''
		%
		\begin{tikzpicture}[remember picture,overlay,,every node/.style={anchor=center}]
		\node[text width=7cm] at (page cs:0.5,0.73){\small \newaddress};
		\end{tikzpicture}

		%%%%%%  The ``opening'' is just the method of address you would like to use at the start of the letter.
		%
		\opening{}

		%%%%%%%%%% Body of letter   %%%%%%%%%%%%%%
		% Remove it if you do not want watermark
		%\watermark{}{}{}

		%%%%%%%%%%% 	Text   %%%%%%%%%%%%%%
		Thank you for the opportunity to submit our manuscript to Nature Energy. Hereby we submit the paper “\hydrogen{} and \carbon{} Network Strategies for the European Energy System”.

		In our study, we introduce the first large-scale investment model that assesses the co-benefits of a carbon-dioxide and a hydrogen transport system in Europe. The model optimizes the production, transport, and storage of various carriers, including electricity, hydrogen, methane, and carbon-dioxide, across all relevant sectors.

		The paper appears at an important time, with policy makers and industry leaders concretizing plans and investments for hydrogen and carbon-dioxide management systems. In particular, it remains unclear how and to what extent synthetic fuel production and permanent carbon storage (carbon sequestration) should be integrated into the future energy system.

		Our analysis not only confirms the crucial flexibility that \hydrogen{} and \carbon{} transport systems bring to \carbon{} emission neutral energy systems, but also illustrates how their combination further enhances energy efficiency and social welfare. We show that the hydrogen grid optimally transports low-cost hydrogen from areas with good renewable resources to regionally-distributed carbon sources such as industrial clusters and biomass combustion plants to enable synthetic fuel production at site. A carbon-dioxide network, on the other hand, is most cost-beneficial to transport low-cost carbon from point sources near the coast to sequestration sites to permanently store it. In such a setup, the system unlocks most carbon capture potentials from biomass and reduces the reliance on cost-intensive direct air capture. Moreover, the benefits of deploying both networks increase when enforcing a net carbon-removal target, leading to system-cost reductions of 6 billion Euros per year.

		Our analysis underlines the importance of incorporating hydrogen and carbon dioxide networks into European energy policy for a resilient, carbon-neutral or negative future energy system.

		We did not have any prior discussions with a Nature Energy editor about this manuscript. The manuscript has not been published and is not under consideration for publication elsewhere.

		We believe that our findings are in line with the scope of Nature Energy and will be of interest to your readership. We want to thank the editor for taking the time to consider the publication of our paper in Nature Energy.

		%%%%%%% ``closing'' sets the sign-off line.
		\closing{Sincerely,}
		% Comment out/in the lines below as necessary
		%\encl{If an enclosure is provided, let them know what it is.}
		%\ps{A postscript if that is a thing you do.}
		%\cc{Someone Who Cares (and is copied).}
	\end{letter}

\end{document}
%
%%-----------------------------------------------------------------------------------------%%