### 2024-09-10 report/submissions/journal/review-responses/review-responses-reviewer-1.md
update colors; insert remaining internal review comments; separate review responses

# Response to Reviewer 1

We thank the reviewers for their helpful comments. We have rerun all scenarios in the paper with improved assumptions as suggested by the authors (improved accounting for leakage and compression losses) as well as new scenarios requested by the reviewers for a sensitivity to sequestration rates and technology costs.

Please let us reply to the reviewers' comments below.

Comment:

> The article presents an interesting analysis of the foreseen H2 and CO2 transport infrastructures.
> Literature is not accurate and there are many references cited without going in their details.
> The study lacks of realistic elements such as leakage issues, location of compression stations in relation with the capture location.

Thank you for your comments. We appreciate your feedback and have addressed your concerns in the revised manuscript.

We have expanded the referenced literature to provide more context for the study, as well as discussing recent policy measures for carbon management in the EU. Since the study touches a wide range of topics in the political and technological landscape, we have tried to find a good balance between providing a comprehensive overview and keeping the text concise.

To simulate a more realistic transport system, the following assumptions have been integrated into the energy system model for transport efficiencies and compression requirements. For hydrogen pipelines an average loss of 1.2% per 1000 km, aligning with industry estimates that account for hydrogen's high permeability and small molecular size, which can lead to higher losses compared to other gases​. The compression energy required is assumed to be 1.8% of the transported energy, which is consistent with the significant energy demands for compressing hydrogen due to its low energy density​ [Environmental Defense Fund​](https://www.edf.org/sites/default/files/documents/Pipeline%20Methane%20Leaks%20Report.pdf). For natural gas pipelines, the model assumes a 1% loss rate per 1000 km, accounting for typical leakage and operational losses in well-maintained natural gas infrastructure​ as reported in [Global Energy Monitor](https://www.gem.wiki/Natural_gas_transmission_leakage_rates). The compression requirement is set at 1%, which falls within the common range for natural gas pipeline operations​ [Environmental Defense Fund​](https://www.edf.org/sites/default/files/documents/Pipeline%20Methane%20Leaks%20Report.pdf)​. For CO2 pipelines an average loss rate of 0.8% per 1000 km is reflecting a generally low leakage rate observed in supercritical or dense phase CO2 transport​ [SpringerLink](https://link.springer.com/article/10.1007/s11356-023-27289-3) [MDPI](https://www.mdpi.com/1996-1073/14/15/4601)​. The compression energy needed is set to 250 kWh per 1000km and tonne CO2 reflecting the energy required to compress and maintain CO2 in a supercritical state for efficient transport​.

Comment:

> There are several graphs that make the reader confused rather than explain the amount of work done by the authors.

We have revised the layout of the figures to make them more accessible to the reader and have adjusted the color scheme where colors tended to be quite similar. However, we would appreciate more specific feedback on which figures you find confusing so that we can address these issues more effectively. Most of the figures contain a lot of information that we feel is necessary to support the argument. We have tried to make them as clear as possible.

Comment:

> The results are not understandable in terms of transferring into policies.

We have written a short policy outlook in the conclusion as well as linking in the introduction to the recently-announced industrial carbon management strategy of the European Commission. Our addition to the conclusion reads:

"For policy makers these results show the need for coordinated planning across sectors like carbon, hydrogen and synthetic fuels since they are strongly interacting, as well as across borders because of the localized nature of good renewable resources and sequestration potentials. Deploying multiple networks offers some robustness should unforeseen problems arise with one of them, and our results show that the system can even cope with neither a CO2 nor H2 network for a cost penalty.  Many of the technologies have not yet been deployed at scale before and will need support for financing, regulation and gaining the acceptance of local populations."